---
title: "homework_3"
output:
  pdf_document: default
  html_document: default
---
#Homework Week 4: Population counts for endangered species


First, we install the required packages: devtools, tidyverse and vroom.
```{r setup, include=FALSE}
install.packages("devtools", dependencies =TRUE)
library("devtools")
install.packages("tidyverse")
library("tidyverse")
install_github("r-lib/vroom", force = TRUE)
library(vroom)
```

We then import both datasets using vroom:
```{r pop}
pop_1 <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%203/to_sort_pop_1.csv")
head(pop_1)
```

```{r pop 2}
pop_2 <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%203/to_sort_pop_2.csv")
head(pop_2)
```

Now, we first add a new column to both population datasets to specify which population they are (1 or 2)
```{r add pop column}
pop_1$Population=1
pop_2$Population =2
```

Then, we can go ahead and use the "merge" function to merge the two data frames, setting the all parameter to TRUE thus including all columns in both data frames.

```{r merge datasets}
pops <- merge(pop_1,pop_2, all = TRUE)
```

Finally, we convert the new merged data frame to long format, specifying the 5 columns we want to keep and merging the other columns with new variables "Date" and "Pop count".
```{r convert to long format}
pops_long <- pops %>%
  
  pivot_longer(cols = -c(species,
                         primary_threat,
                         secondary_threat,
                         tertiary_threat,
                         Population),
               names_to = "Date",
               values_to = "Pop count")

head(pops_long)
```